---
title: "COVID-19 Analysis"
description: |
  Stay home, or become [Italy](http://www.cidrap.umn.edu/news-perspective/2020/03/doctors-covid-19-pushing-italian-icus-toward-collapse)
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
install.load::install_load(c('distill','dplyr', 'tidyverse', 'ggplot2', 'plotly', 'rmarkdown', 'magrittr', 'RCurl', 'lubridate', 'DT', 'ggthemes'))
```

```{r}
fp_data = 'https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv'

# Read in data
dfc = read_csv(fp_data) %>%
  select(-Lat, -Long) %>%
  gather(date_raw, confirmed, 
         -`Province/State`, -`Country/Region`) %>%
  mutate(date = mdy(date_raw))

# Clean column names
colnames(dfc) %<>% 
  tolower() %>% 
  str_replace_all(fixed('/'), '_') %>%
  str_replace_all(fixed(' '), '_')

ggplot2::theme_set(theme_minimal())

# US
dfc %>%
  filter(country_region == 'US') %>%
  group_by(date) %>%
  summarize(confirmed = sum(confirmed)) %>% 
  ggplot(aes(x = date, y = confirmed)) +
  geom_line() +
  ggtitle('US Cases are on an Exponential Rise')

```
If we don't learn the lessons other countries have, we'll repeat their recently-learned history. Staying home saves lives.

## The US Had a Delayed Start
While China was the first to get hit by COVID-19, everyone else is actually on a similar growth trajectory (unless we take action #stayhome!). The US was in a lull for many weeks, probably because access to China was halted. But then cases came in from other sources. Now, we're past containment.

This graph shows how many confirmed cases there have been in the world relative to days since the country first observed an onset. 

```{r}
# Day since 1
df_sum <- dfc %>%
  group_by(country_region, date) %>%
  summarize(confirmed = sum(confirmed)) %>%  
  arrange(country_region, date)

df_tmp <- df_sum %>%
  filter(confirmed > 0) %>%
  mutate(day_since_start = row_number()) %>%
  ungroup() %>%
  #filter(country_region %in% c("China", 'US', 'Italy', 'France', 'Iran')) %>% #View
  highlight_key(~country_region) 
  
p <- ggplot(df_tmp, aes(x=day_since_start, 
                        y=confirmed, 
                        color=country_region)) +
  geom_line() +
  labs(title='Total Confirmed COVID-19 Cases by Days Since First Onset',
          x = 'Days Since Start of Onset',
          y = 'Total Confirmed') + 
  theme(legend.position = "none")

ggplotly(p) %>% 
  highlight('plotly_hover', 
          selectize = T, 
          # dynamic=T,
          defaultValues = c('China', 'Italy', 'US'))
```

## COVID-19 Doubles Really Fast, for Everyone
No matter how long it takes a country to hit 500 cases, they all double at the same rate.

Italy hit it's 500th case on day 28 of being exposed. It then doubled in 2 days. Then again. Then again.

The US didn't hit it's 500th case until day 47 of the virus being in the US. This was 47 days of a false sense of security. This was 47 days we all watched China double exponentially. Then *we* started doubling.

There are eight countries that have hit 4,000 cases as of March 16, 2020. It takes about 1-3 days for the virus to double, regardless of the country it's been present in.

```{r}
# How many countries are there:
df_firsts <- df_sum %>%
  group_by(country_region) %>%
  filter(confirmed >= 1) %>%
  mutate(max_confirmed = max(confirmed)) %>%
  arrange(country_region, date) %>%
  mutate(days_since_first = row_number()) 

df_firsts %>%
  filter(confirmed > 500) %>%
  mutate(
    days_until_500 = min(days_since_first),
  ) %>%
  filter(confirmed > 1000) %>%
  mutate(
    days_until_1k = min(days_since_first),
    days_500_to_1k = days_until_1k - days_until_500
  ) %>% 
  filter(confirmed > 2000) %>%
  mutate(
    days_until_2k = min(days_since_first),
    days_1k_to_2k = days_until_2k - days_until_1k
  ) %>% 
  filter(confirmed > 4000) %>%
  mutate(
    days_until_4k = min(days_since_first),
    days_2k_to_4k = days_until_4k - days_until_2k
  ) %>% 
  filter(date == Sys.Date() - 1) %>%
  select(country_region, confirmed, days_until_500, days_500_to_1k, days_1k_to_2k, days_2k_to_4k, days_since_first) %>%
  arrange(desc(days_until_500)) %>%
  datatable(
    colnames = c(
      'Country', 
      'Total Confirmed',
      'Days to Hit 500', 
      'From 500 to 1k',
      'From 1k to 2k',
      'From 2k to 4k',
      'Total Days Since First'
    ),
    options = list(
      pageLength = 100,
      dom = 't'
    )
  )
```

## Flatten the Curve
Looking at countries that have had 500 cases, we can see how the US is in alignment with other countries, especially Italy.

```{r}
df_tmp <- df_sum %>%
  filter(confirmed > 500) %>%
  mutate(day_since_500 = row_number()) %>%
  filter(country_region != 'Cruise Ship') %>%
  ungroup() %>%
  #filter(country_region %in% c("China", 'US', 'Italy', 'France', 'Iran')) %>% #View
  highlight_key(~country_region) 
  
p <- ggplot(df_tmp, aes(x=day_since_500, 
                        y=confirmed, 
                        color=country_region)) +
  geom_line() +
  labs(title='Total Confirmed COVID-19 Cases by Days \nSince First Reaching 500',
          x = 'Days Since 500 Cases',
          y = 'Total Confirmed') +
  theme(legend.position = "none")

ggplotly(p) %>% 
  highlight('plotly_click', 
          selectize = T, 
          defaultValues = c('Italy', 'US'))
```

To hone in on the trends, we see that the US is just ten days behind Italy. Read the headlines. If we don't take action, we'll be in the same position.

```{r}
df_sum %>%
  filter(confirmed > 100) %>%
  mutate(days_since = row_number(),
         days_since = ifelse(country_region == 'US', days_since - 1 , days_since)) %>%
  filter(country_region %in% c('Italy', 'US')) %>%
  ggplot(aes(x = days_since, y = confirmed, fill = country_region)) +
  geom_bar(stat = 'identity', position='dodge') +
  scale_fill_fivethirtyeight(name='Country') + 
  labs(title='The US is Just Ten Days Behind Italy',
       x = 'Days Since Hitting 100 Confirmed Cases',
       y = 'Total Confirmed Cases',
       subtitle = 'US-confirmed cases shifted one day to the left to align trends',
       caption = 'Data Source: github.com/CSSEGISandData/COVID-19\nbryanwhiting.com')
```


